---
title: "Class Size"
author: "Tanay Janmanchi"
format: html
execute:
  echo: FALSE
---

```{r}
#| label: setup
#| message: FALSE

library(tidyverse)
library(brms)
library(tidybayes)
library(gtsummary)
```

```{r}
#| label: reading-data
#| message: FALSE

class <- read_csv("Data/STAR.csv", show_col_types = FALSE)
```

```{r}
#| label: cleaning
#| cache: TRUE

class <- class |> 
  mutate(kinder = recode(classtype,
                         `1` = "small",
                         `2` = "regular",
                         `3` = "regular with aid")) |> 
  mutate(race = recode(race,
                       `1` = "white",
                       `2` = "black",
                       `3` = "hispanic",
                       `4` = "hispanic",
                       `5` = "others",
                       `6` = "others")) |>
  filter(classtype != 3) |>
  select(g4math, 
         kinder) |>
  drop_na()

fit_class <- brm(formula = g4math ~ kinder,
                 data = class,
                  family = gaussian(),
                  silent = 2,
                  refresh = 0)
```

```{r}
#| label: models
#| fig-height: 7
#| fig-width: 15

ndata = tibble(kinder = c("regular", "small"))

fit_class |>
  add_epred_draws(newdata = ndata) |>
  ggplot(aes(x = .epred)) +
  geom_density(aes(fill = kinder),
                   alpha = 0.5) +
  scale_y_continuous(label = NULL) +
  theme(axis.ticks.y = element_blank(),
        text = element_text(size = 20)) +
  labs(title = "Posterior for Class Score",
       x = "Exam Score",
       subtitle = "Class Size has no effect on exam score",
       y = NULL,
       fill = "Class Size")
# 
# class |>
#   add_epred_draws(newdata = tibble(kinder = c("regular", "small"))) |>
#   ungroup() |>
#   group_by(.draw) |>
#   select(kinder, .epred, .draw) |>
#   pivot_wider(names_from = kinder, values_from = .epred) |>
#   mutate(causal = small - regular) |>
#   ggplot(aes(x = causal)) +
#   geom_density()

```

Using data from Tennessee Project STAR, we seek to understand the causal effect of putting students in smaller classrooms on their grade 4 math scores. Students are randomly assigned to a Kindergarten class of size small or regular, then in grade 4, the students take a math test to measure their progress. One challenge we encountered was that the method of evaluating grade 4 math score might not have been consistent over the years. We found that there is little to no effect of class size on math exam scores. We are 95% confident that the causal effect is between -4.5 and 3.9 and predict the causal effect to be -0.36.




$$ score_i = \beta_0 + \beta_1 class\_size $$

```{r}
#| label: courage
# fit_class
# fixef(fit_class)
# pp_check(fit_class)
tbl_regression(fit_class)
```

<!-- **Question**: What is the causal effect of class size on childrens' math score? -->

<!-- **QoI**: Grade 4 math score -->

<!-- **Preceptor Table**: One row for each individual student, two potential outcomes: math score for regular class size and math score for smaller class size. -->

<!-- **Covariates**: Race, Years of small class size -->

<!-- **Treatment**: Class size -->

<!-- **Moment in Time**: 2024 -->

<!-- **Validity**: Method of evaluating grade 4 math score might not have been consistent over the years. -->

<!-- **Stability**: Children from the 1990s may have had a preference in class size as opposed to now. -->

<!-- **Representativeness**: Children in Tennessee might not represent the greater population. -->

<!-- **Unconfoundedness**:  -->



